---
title: "Missing Data Analysis"
author: "William Murrah"
date: ''
output:
  html_document:
    fig_height: 3
    fig_width: 5
  pdf_document:
    fig_height: 3
    fig_width: 5
  word_document:
    fig_height: 3
    fig_width: 5
---
```{r, include=FALSE}
# Don't delete this chunk if you are using the mosaic package
# This loads the mosaic and dplyr packages
require(mosaic)
```

```{r, include=FALSE}
# Some customization.  You can alter or delete as desired (if you know what you are doing).

# This changes the default colors in lattice plots.
trellis.par.set(theme=theme.mosaic())  

# knitr settings to control how R chunks work.
require(knitr)
opts_knit$set(root.dir = "../../")
opts_chunk$set(
  message = FALSE,
  comment = NULL,
  tidy=FALSE,     # display code as typed
  size="small"    # slightly smaller font for code
)
```

## Packages we will use:
```{r}
require(texreg)  
require(mice)
require(VIM)
require(ztable)

```

```{r}
doctype <- 
   # "latex"
   "html"
txreg <- ifelse(doctype == "latex", texreg, htmlreg)
options(ztable.type = doctype)
```

## Notation we will use:

* $n$ = number of units (number of cases or people) indexed by $i$.
* $p$ = number of variables (including outcome and predictors), indexed by $j$.
* $Y$ = $n\times p$ matrix containing the data values on  $n$ units for $p$ variables in the sample.
* $R$ = response indicator, a $n\times p$ matrix with each cell containing either a `0` or a `1`, where

$$
r_{ij} = 
\begin{cases}
  \text{1 if $y_{ij}$ is observed and,} \\
  \text{0 if $y_{ij}$ is missing.}
\end{cases} \tag{1}
$$

* $Y_{obs}$ = the observed data, collectively (i.e. contains all elements $y_{ij}$ where $r_{ij} = 1$).
* $Y_{mis}$ = the missing data, collectively (i.e. contains all elements $y_{ij}$ where $r_{ij} = 0$). 


```{r , echo=FALSE}
### define data sets pattern1-pattern4 with four patterns
set.seed(1234)
data <- matrix(data = sample(x = 1:100, 
                             size = 4*8*3, 
                             replace=TRUE), 
               nrow = 8*4,
               dimnames = list(NULL, 
                             c("A","B","C")))
data <- as.data.frame(data)
data[c(31:32),"A"] <- NA
data[c(15:16,22:24,30:32),"B"] <- NA
data[c(6:8,12:16,17:21,27:29),"C"] <- NA

mdpat <- cbind(expand.grid(rec = 8:1, pat = 1:4, var = 1:3), r=as.numeric(as.vector(is.na(data))))
mdpat2 <- 1 - is.na(data)
pattern1 <- data[1:8,]
pattern2 <- data[9:16,]
pattern3 <- data[17:24,]
pattern4 <- data[25:32,]
### Figure 4.1
types <-  c("Univariate","Monotone","File matching","General")
tp41 <- levelplot(r~var+rec|as.factor(pat), data=mdpat, 
                  as.table=TRUE, aspect="iso",
                  shrink=c(0.9), 
                  col.regions = mdc(1:2),
                  colorkey=FALSE,
                  scales=list(draw=FALSE),
                  xlab="", ylab="",
                  between = list(x=1,y=0),
                  strip = strip.custom(bg = "grey95", style = 1,
                                       factor.levels = types))
general <- pattern4
rownames(general) <- 1:8
```

\newpage

## Missing Data Patterns
```{r , echo = FALSE, fig.width=8, fig.height=5}

print(tp41)
```

```{r}

print(tp41[4])
```

## A simple data frame with some missing data:
```{r}
general
```

### Create an $R$ matrix
```{r}
R <- 1 - is.na(general)
R
```

```{r}
md.pattern(pattern4)

```

```{r}
aggr(general, numbers = TRUE, prop = FALSE)
```

## Proportion of usable cases

Imputing $Y_j$ from $Y_k$, the *proportion of usable cases* is the number of cases missing in $Y_j$ that are observed in $Y_k$ divided by the number of missing cases in $Y_j$.

```{r}
p <- md.pairs(general)
p$mr/(p$mr + p$mm) # proportion of usable cases.
```

## Influx and Outflux

Influx - $I_j$ the *influx coefficient* is how well the other variables connect to $Y_j$.

* $I_j$ is `0` for a completely observed variable
* $I_j$ is `1` for a completely missing variable

Outflux - $O_j$ the *outflux coefficient* is how well $Y_j$ is connected to other variables.

* $O_j$ is `1` for a completely observed variable
* $O_j$ is `0` for a completely missing variable

```{r, warning=FALSE, fig.width=4, fig.height =4, fig.show = 'hold'}
print(tp41[1])
fluxplot(pattern1, main = NULL)
```

```{r, warning=FALSE, fig.width=4, fig.height =4, fig.show = 'hold'}
print(tp41[2])
fluxplot(pattern2, main = NULL)
```

```{r, warning=FALSE, fig.width=4, fig.height =4, fig.show = 'hold'}
print(tp41[3])
fluxplot(pattern3, main = NULL)
```

```{r, warning=FALSE, fig.width=4, fig.height =4, fig.show = 'hold'}
print(tp41[4])
fluxplot(general, main = NULL)
```

\newpage

# Missing Data Analysis


## Data we will use:

```{r, results='asis'}
#  ?nhanes
data("nhanes2")
ztable(nhanes)

```

## Model

$$
\text{chl}_i = \beta_0 + \beta_1 \text{age}_i + \beta_2 \text{bmi}_i + \beta_3 \text{hyp}_i + \epsilon_i \tag{2}
$$

## Missing Data Patterns
```{r}
md.pattern(nhanes)
aggr(nhanes, numbers = TRUE, prop = FALSE)
```

```{r}
p <- md.pairs(nhanes)
p$mr/(p$mr + p$mm)
```

```{r , fig.height=6, fig.width=6}
fluxplot(nhanes)
```




## Predictors of missingness

```{r}
round(x = cor(x = nhanes, 
              use = "pairwise.complete.obs"), 
      digits = 2)
```

```{r, results='asis'}
pmA <- glm(is.na(chl) ~ age + bmi + hyp, family = binomial, data = nhanes)
pmB <- glm(is.na(bmi) ~ age + chl + hyp, family = binomial, data = nhanes)
pmC <- glm(is.na(hyp) ~ age + bmi + chl, family = binomial, data = nhanes)
txreg(list(pmA, pmB, pmC), custom.model.names = c("chl", "bmi", "hyp"))
```


```{r, echo=FALSE}
# achieve <- read.csv('data/Achieve.csv')
# achieve <- achieve[ ,c("geread", "gevocab", "gender", "age")]
# achieve$gender <- achieve$gender - 1
# achieve$age <- achieve$age/12

```

```{r, echo=FALSE}
# mod <- lm(geread ~ gevocab + gender + age, data = achieve)
# summary(mod)
```

```{r, echo=FALSE}
# b0 <-  1.97
# b1 <-  0.53
# b2 <-  0.04
# n <- 10320
# sigma <- 1.97
# 
# set.seed(123)
# gevocab <- rnorm(n = n, mean = 4.94, sd = 2.37)
# gender <- rep(0:1, n/2)
# 
# geread <- b0 + b1*gevocab + b2*gender + rnorm(n = n, mean = 0, sd = sigma )
# simdata <- data.frame(read = geread, 
#                       vocab = gevocab,
#                       female = factor(gender, 
#                                       labels = c("male", "female")))
```

```{r, echo=FALSE}
# logistic <- function(x) exp(x)/(1 + exp(x))
# inv.logit()
# set.seed(1234)
# r.mcar <- 1 - rbinom(n, 1, 0.50)
# r.mar <- 1 - rbinom(n, 1, logistic(simdata$vocab))
# set.seed(1234)
# r.mnar <- 1 - rbinom(n, 1, inv.logit(simdata$read))
```

```{r echo=FALSE}
sessionInfo()  # could use devtools::session_info() if you prefer that
```
  