---
title: "Analsis of Imputed Data"
author: "William Murrah"
date: ''
output:
  html_document:
    fig_height: 6
    fig_width: 6
  pdf_document:
    fig_height: 6
    fig_width: 6
  word_document:
    fig_height: 6
    fig_width: 6
bibliography: ref.bib    
---
```{r, include=FALSE}
# Don't delete this chunk if you are using the mosaic package
# This loads the mosaic and dplyr packages
require(mosaic)
```

```{r, include=FALSE}
# Some customization.  You can alter or delete as desired (if you know what you are doing).

# This changes the default colors in lattice plots.
trellis.par.set(theme=theme.mosaic())  

# knitr settings to control how R chunks work.
require(knitr)
opts_knit$set(root.dir = "../../")
opts_chunk$set(
  message = FALSE,
  comment = NULL,
  tidy=FALSE,     # display code as typed
  size="small"    # slightly smaller font for code
)
```

## Packages we will use:
```{r}
require(texreg)  
require(mice)
require(VIM)
require(ztable)
require(lattice)
require(mice)
require(psych)
require(miceadds)
source("~/Rfunc/miralm.R")
```

```{r}
doctype <- 
   # "latex"
   "html"
txreg <- ifelse(doctype == "latex", texreg, htmlreg)
options(ztable.type = doctype)
```


## Data example

NHANES

```{r}

headTail(nhanes, tlength = 10) # from `psych` package
md.pattern(nhanes)
lmfit <- lm(chl~bmi+age,data=nhanes)
```

```{r, results='asis'}
htmlreg(lmfit, custom.model.names = "listwise")
```



```{r, cache=TRUE}
imp <- mice(data = nhanes, m = 5, maxit = 20, seed = 20151110, 
            printFlag = FALSE)
```


```{r}
plot(imp)
stripplot(imp)
densityplot(imp)
```


## What **NOT** to do with imputed data

1. Average data sets

2. Stack data sets
```{r, results='asis'}
stacked <- complete(imp, "long")
stackfit <- lm(chl~bmi+age,data=stacked)
htmlreg(list(lmfit, stackfit), custom.model.names = c("listwise", "stacked"))
```

## Repeated Analyses: What you **SHOULD** do

The appropriate way to analyze MI data is to do repeated analyses and pool parameter estimates.

The following block quote is from @molenberghs2014handbook.

Multiple imputation is a procedure that enables the user to carry on with their standard
analysis almost as if the dataset contained no missing values. The method proceeds as
follows:

> 1. Create a number ($m$) of copies of the incomplete dataset, and use an appropriate
procedure to impute (fill in) the missing values in each of these copies. Since we do
not know the true values that are missing it seems reasonable that the imputed values
used in each copy should in general be different from each other.

> 2. For each completed (imputed) copy of the dataset, carry out the standard analysis
that would have been performed in the absence of missing values, and store the parameter
estimates of interest, along with their estimated standard errors (or variance/covariance
matrix in the case of a multivariate parameter of interest). Here ... the focus will be on a single (univariate) parameter of interest,
which we denote $\beta$. The estimate of $\beta$ obtained from the $k^{th}(k = 1, . . .,m)$ completed
dataset is denoted $\beta^{(k)}$ and its (estimated) variance $V^{(k)}$. (Note that this step implicitly
assumes that the inference desired from the analysis can be effectively reduced
to a parameter estimate and standard error.)

> 3. Use a pair of formulas widely known as "Rubin's rules,"" firstly to create a combined
estimate of the parameter, as the average of the $m$ separate estimates:

$$
\hat{\beta}^{MI} = \frac{1}{m} \sum_1^m \hat{\beta}^{(K)},
$$

> and then to obtain a standard error for this estimate, as the square root of  the following combined variance estimate:

$$
V^{(MI)} = \bar{V} + \left(1 + \frac{1}{m} \right)B,
$$

> where $\bar{V} = \sum_1^m V^{(k)}/m$, and $B = \sum_1^m (\hat{\beta}^{k} - \hat{\beta}^{MI})^2/(m - 1)$, which is an estimate of the *between-imputation* variance of the parameter of interest. In general multiple-imputation inference proceeds in the usual way from these results, by forming test statistics and confidence intervals under the assumption that $(\hat{\beta}^{MI} - \beta)/\sqrt{V^{MI}}$ follows either a standard normal or $t$ distribution.

### Repeated Analyses:
```{r}
fit <- with(imp, lm(chl ~ bmi + age))
# Grab estimates from first analysis:
coef(fit$analyses[[1]])

# Grab estimates from second analysis:
coef(fit$analyses[[2]])
```

### Pool estimates:
```{r}
est <- pool(fit)
summary(est)
```

### Using other functions on imputed data:
```{r}
expr <- expression(mean.bmi <- mean(bmi))
out <- with(imp, eval(expr))
unlist(out$analyses)
mean(unlist(out$analyses))
```


```{r, eval = FALSE}
mids2mplus(imp, path = "data/mplusData/")
```

```{r, eval=FALSE}
mids2spss(imp, path = "data/spssData/")
```

```{r, results = 'asis'}
source("~/FOCAL/Rpackages/cpar/R/miralm.R")
miralmObj <- miralm(mira = fit, n.obs = 25)
htmlreg(list(lmfit, stackfit, miralmObj), custom.model.names = c("listwise", "stacked", "mice"))
```

```{r echo=FALSE}
sessionInfo()  # could use devtools::session_info() if you prefer that
```
  
# References  
  