---
title: "Basic Simulation in R"
author: "William Murrah"
date: ""
output: 
  html_document:
    fig_height: 3
    fig_width: 5
  pdf_document:
    fig_height: 3
    fig_width: 5
  word_document:
    fig_height: 3
    fig_width: 5
---


```{r, include=FALSE}
# Don't delete this chunk if you are using the mosaic package
# This loads the mosaic and dplyr packages
require(mosaic)

```

```{r, include=FALSE}
# Some customization.  You can alter or delete as desired (if you know what you are doing).

# This changes the default colors in lattice plots.
trellis.par.set(theme=theme.mosaic())  

# knitr settings to control how R chunks work.
require(knitr)
opts_knit$set(root.dir = "../../")
opts_chunk$set(
  comment = NULL,
  tidy=FALSE,     # display code as typed
  size="small"    # slightly smaller font for code
)
# This loads the mosaic data sets.  (Could be deleted if you are not using them.)
library(texreg)
library(stargazer)               
```

## Why do I need to know how to simulate data?

* It is closely related to several statistical procedures (e.g. bootstraping, MCMC)
* It is closely related to multiple imputation
* It is helpful in checking the assumptions of your model

## Probability Distributions

In R, enter the following code into the console:

```{r, eval=FALSE}
?distributions
```

### Some of the distributions available in R

 R code  | distribution
 --------|-------------
`unif`   | uniform
`norm`   | normal
`binom`  | binomial 
`chisq`  | chi-squared
`t`      | Student's t


### Distribution prefixes

prefix  |  function
--------|-----------
`d`     | density or probability mass function (pmf)
`p`     | cumulative distribution function (cdf)
`q`     | quantiles
`r`     | random number generation (actually peudo-random) 


```{r} 
set.seed(082184) # Set the seed for reproducible results

```

```{r}
x <- runif(n = 1000,
           min =-3, 
           max = 3)
x[1:10]

hist(x, 
     col = 'skyblue', 
     xlim=c(-5,5),
     main = "1000 random samples from uniform distribution")
```

```{r}
x <- rnorm(n = 1000, 
           mean = 0, 
           sd = 3)
x[1:10]

hist(x, 
     col = 'skyblue', 
     main = "1000 random samples from normal distribution")
```

## Simple Simulation of Linear Regression Model


```{r}
achieve <- read.csv('data/Achieve.csv')
achieve <- achieve[ ,c("geread", "gevocab", "gender")]
achieve$gender <- achieve$gender - 1
stargazer(achieve, type = 'text', column.sep.width = '25pt')
```


```{r , results='asis'}
mod <- lm(geread ~ gevocab + gender, data = achieve)
htmlreg(mod)
```

### Model equation
$$
y = \beta_0 + \beta_1 X_1 + \beta_2 X_2 + \epsilon \tag{1}
$$

$$
y = 1.95 + .53 \textrm{gevocab} + .03 \textrm{gender} + \epsilon \tag{2}
$$

$$
\epsilon = N(0, \sigma) \tag{3}
$$

```{r, results='asis'}
b0 <- 1.95
b1 <- .53
b2 <- .03
n <- 10320
sigma <- 1.97

set.seed(123)
gevocab <- rnorm(n = n, mean = 4.94, sd = 2.37)

gender <- rep(0:1, n/2)

geread <- b0 + b1*gevocab + b2*gender + rnorm(n = n, mean = 0, sd = sigma )

sim.mod <- lm(geread ~ gevocab + gender)
htmlreg(list(mod, sim.mod), custom.model.names = c("original", "simulated"))
```


```{r, fig.height=8, fig.width= 16}
par(mfrow = c(1,2))
plot(geread ~ gevocab, data = achieve, main = "original")
plot(geread ~ gevocab, main = "simulated")
cor(achieve)
cor(geread, gevocab)
range(geread)
```

### Session Information

  
```{r echo=FALSE}
sessionInfo()  # could use devtools::session_info() if you prefer that
```
  