---
title: "Imputation in Practice"
author: "William Murrah"
date: ''
output:
  html_document:
    fig_height: 6
    fig_width: 6
  pdf_document:
    fig_height: 6
    fig_width: 6
  word_document:
    fig_height: 6
    fig_width: 6
---
```{r, include=FALSE}
# Don't delete this chunk if you are using the mosaic package
# This loads the mosaic and dplyr packages
require(mosaic)
```

```{r, include=FALSE}
# Some customization.  You can alter or delete as desired (if you know what you are doing).

# This changes the default colors in lattice plots.
trellis.par.set(theme=theme.mosaic())  

# knitr settings to control how R chunks work.
require(knitr)
opts_knit$set(root.dir = "../../")
opts_chunk$set(
  message = FALSE,
  comment = NULL,
  tidy=FALSE,     # display code as typed
  size="small"    # slightly smaller font for code
)
```

## Packages we will use:
```{r}
require(texreg)  
require(lattice)
require(mice)
require(VIM)
require(ztable)
require(mice)
require(psych)
source("~/Rfunc/miralm.R")
```

```{r}
doctype <- 
   # "latex"
   "html"
txreg <- ifelse(doctype == "latex", texreg, htmlreg)
options(ztable.type = doctype)
```


## Data example

NHANES

```{r}
data <- nhanes
headTail(data, tlength = 10) # from `psych` package
```



```{r, eval = FALSE}

## DON'T DO THIS!!
imp <- mice(nhanes)
```


### Seven steps:
1. Ignorable or nonignorable?
2. Model form
3. Predictors
4. Derived variables
5. Imputation order 
6. Starting values and number of iterations
7. Number of imputations


```{r, eval = FALSE}
imp <- mice(data = nhanes,        # data
            m = 5,               # number of imputations
            method = ,            # vector of imputation methods 
                                  #  for each variable
            predictorMatrix = ,   #
            visitSequence = ,     # order of imputations
            maxit = 5,             # number of iterations
            seed = ,              # random number seed
            data.init =           # data frame with starting values
              )
```

```{r}
imp0 <- mice(data = nhanes, maxit = 0)
predMat <- imp0$pred
predMat

imp0$method

meth <- imp0$method
meth[3] <- "norm"
meth
```

```{r}
imp <- mice(data = nhanes, m = 5, maxit = 10, seed = 20151110)
str(imp)
```


## Diagnostics

### Trace plots

*Diagnosing Convergence* 

1. Different chains should mix well (freely intermingle)
2. Chains should not show trends
3. Variance between chains is no larger than variance within chains.

#### Poor or No convergence

```{r, echo = FALSE, cache = TRUE}
ini <- mice(boys,max=0,print=FALSE)
meth <- ini$meth
meth["bmi"] <- "~I(wgt/(hgt/100)^2)"
imp.bmi1 <- mice(boys, meth=meth, maxit=20, seed=60109, printFlag = FALSE)

### Figure 5.7
### Does not reproduce exactly, but the message is the same
plot(imp.bmi1,c("hgt","wgt","bmi"), col=mdc(5), lty=1:5)
```

#### Healthy convergence

```{r, echo = FALSE, cache = TRUE}
pred <- ini$pred
pred[c("hgt","wgt"),"bmi"] <- 0
imp.bmi2 <- mice(boys, meth=meth, pred=pred, maxit=20, seed=60109, 
                 printFlag = FALSE)
plot(imp.bmi2,c("hgt","wgt","bmi"), col=mdc(5), lty=1:5)
```

### Stripplots

```{r}
stripplot(imp.bmi2)
```


### Density plots

```{r}
densityplot(imp.bmi2)
```

```{r echo=FALSE}
sessionInfo()  # could use devtools::session_info() if you prefer that
```
  